Light inextensible strings under tension in the Schwarzschild geometry

### 2019-08-28 string_physics_teacher.tex
further work on radius, there is no such thing as "the" radius of curvature, only comparison with other curved paths (such as null geodesics)

\noindent which would give~$R$ as a function of~$\left(r,r'\right)$.
Considering~$r'(\phi)=0$ shows that if a string is tangential at a
point with~$r_0>2M$, then it has a radius of curvature of~$r_0^2/2M$
there.  Tangential null geodesics have a radius of curvature of
$r_0^2/3M$, strictly less than that of a string.  This shows that a
taut string would {\em look} curved to a local observer, having a
(visual) radius of curvature of $r_0^2/6M$.  At the photon sphere of a
solar mass black hole, such a string would appear to have a radius of
about $2\unit{km}$, a readily understandable result.

@@

\begin{figure}[p] % Created by "light_closest_approach.R"
\centering
\includegraphics[width=170mm]{light_closest_approach.pdf}
\caption{Null geodesics in the Schwarzschild geometry, tangential
  at~$\phi=0$.  Note the differences between these curves and the taut
  strings shown elsewhere: unlike taut strings, null geodesics have
  positive curvature everywhere, may cross the event horizon inwards,
  and are never tangential to the event horizon}
\label{light_closest_approach}
\end{figure}

>>>>

\noindent which would give~$R$ as a function of~$\left(r,r'\right)$.
Considering~$r'(\phi)=0$ shows that if a string is tangential at a
point with~$r_0>2M$, then it obeys, to first order, $r''=r-2M$, while
tangential null geodesics follow $r''=r-3M$.  This shows that a taut
string would {\em look} curved to a local observer, having a (visual)
radius of curvature of $r_0^2/M\sqrt{1-2M/r_0}$ (the extra term comes
from the fact that appearance is driven by local $ds$, not
Schwarzschild $r$).  At the photon sphere $r_0=3M$, such a string
would appear to have a radius of~$\sqrt{27}M$.  For a solar mass black
hole this would be about $15\unit{km}$, a readily understandable
result.

@@

\begin{figure}[p] % Created by "light_closest_approach.R"
\centering
\includegraphics[width=170mm]{light_closest_approach.pdf}

\caption{Null geodesics in the Schwarzschild geometry, tangential
  at~$\phi=0$.  Note the differences between these curves and the taut
  strings shown elsewhere: unlike taut strings, null geodesics have
  positive curvature everywhere, may cross the event horizon inwards,
  and are never tangential to the event horizon}
\label{light_closest_approach}
\end{figure}
